---
title: "Update_AllWells"
author: "Nick Triozzi"
date: "2022-12-18"
output: html_document
---

## Overview
This code can be used to update a CSV file containing AllWells_updated ArcGIS Online resource with current CalGEM AllWells_gis data. CalGEM updates the dataset daily. AllWells_updated has additional fields that are not included with the CalGEM resource. This procedure will merge CalGEM's data with the AllWells_updated data, retaining fields of interest in the latter. After this procedure is completed, the user must re-publish the AllWells_updated dataset on ArcGIS Online and update any layers or views that used it so that Web Maps will reflect the current status of California wells.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/using.R")
using("esri2sf", "sp", "tidyverse", "rgdal")
```

## Retrieve Data

The GIS datasets used here include AllWells_updated (A CREW resource) and CalGEM's AllWells_gis dataset which is updated daily. Both of these datasets are hosted as Web Feature Services on ArcGIS Online. The CREW resource must first be downloaded as a CSV file and saved in the */data*  directory. CalGEM's WFS can be accessed using a url. 

Below we set variable CalGEM.url to the WFS data and retrieve the data from it's online service.

```{r CalGEM wells}
#-- set url to WFS to variable 
# crew.url = "https://services1.arcgis.com/4TXrdeWh0RyCqPgB/arcgis/rest/services/AllWells_updated/FeatureServer/0"
CalGEM.url = "https://gis.conservation.ca.gov/server/rest/services/WellSTAR/Wells/MapServer/0"

#-- use esri2sf to retrieve data as a sf object
# crew = esri2sf(crew.url) #-- call url
calgem = esri2sf(CalGEM.url) #-- call url

#-- inspect calgem
class(calgem)
str(calgem) # note SpudDate field is character type

```

## Download AllWells_updated

Since this layer is not shared publically it cannot be accessed using a url. The data must be downloaded from the layer overview page <https://ucsb.maps.arcgis.com/home/item.html?id=b7b01223c2d042a6a04e3e6bdcb0a763>. You will need to sign into ArcGIS Online and you will need to be a member of the CREW Group to access and export the layer. Export the layer as a CSV file then download it and save the zip file to the *CREW_R/data* directory. 

Set the variable *well.csv* to the zip file name.

```{r AllWells_updated}
#-- run if the zipped csv file is the only file in the data directory
well.zip = dir("./data", full.names = T)
unzip(well.zip, exdir = "./data") 
#-- Otherwise, specify the zipped filename here (and uncomment)
# well.csv = "AllWells_updated_csv122022.zip"
# unzip(paste0("./data/", well.csv), exdir = "./data") # uncomment with line above to unzip file

#-- read in csv file
well.csv=list.files("./data", pattern="csv", full.names = T)
crew.wells = read.csv(well.csv,header = T)

#-- inspect files
str(crew.wells)
str(calgem)

#-- note changes in well status
status = left_join(data.frame(crew.wells %>% group_by(WllStts) %>% summarise(CREWcount=n())),
          data.frame(calgem %>% group_by(WellStatus) %>% summarise(calGEMcount=n())),
          by=c("WllStts"="WellStatus")
          )
status$diff = status$calGEMcount-status$CREWcount
status[,-4] # shows how many wells changed status in each category
```

## Drop fields from All Wells

## Join tables based on API

Now that we have both datasets in memory we will join the tables together, matching records based on API. For the CalGEM dataset, API is stored as a character. In AllWells_updated it is an integer. We will need to reformat API in the CalGEM dataset to integer so that we can update the file on ArcGIS Online.


```{r, join tables}
#-- note different data types in API columns
str(crew.wells$API)
str(calgem$API)

#-- set API to integer
calgem$API = as.integer(calgem$API)
#-- set Section to integer
calgem$Section = as.integer(calgem$Section)

#-- join tables
new.wells = full_join(crew.wells, calgem, by=c("API"="API"))

#-- inspect
str(new.wells)

#-- check new table has new records
isTRUE(nrow(new.wells)==nrow(calgem))
```

## Reformat Join Table

The datafile used to upate the ArcGIS Online layer must have the same schema, meaning field types and names in the new table we are creating here must match the existing layer's schema. We will preserve the existing column names and use them later.

```{r format}
#-- existing column names
crew.cols = colnames(crew.wells)
crew.cols = crew.cols[c(-38,-39)]

#-- drop old columns, keep CalGEM columns, re-order to match existing schema and to rename columns
colnames(new.wells[,c(40,2,41:46,48,47,49:56,19,20,57,63,64,58:60,62,61,29:37)]) # columns from CalGEM and CREW files being kept
new.wells.fmt = new.wells[,c(40,2,41:46,48,47,49:56,19,20,57,63,64,58:60,62,61,29:37)] # drop cols and create new table

#-- compare column names. 
data.frame(crew.col.names=crew.cols, new.wells.col.names = colnames(new.wells.fmt))

#-- replace column names
colnames(new.wells.fmt) = crew.cols

#-- function to reformat date fields
to.Date = function(x){
  for (i in 1:length(x)){
    if(is.na(x[i])==FALSE){
      x[i]=paste(gsub("-", "/", x[i] ), "12:00:00 PM")
    }
  }
  return(x)
}
#-- Replace SpudDate with formatted dates
new.wells.fmt$SpudDate = to.Date(new.wells.fmt$SpudDate)


```

## Prepare for upload as shapefile
Now we will create a shapefile from the updated data. The shapefile consists of several files so we will make a new directory to store the files. The directory needs to be zipped before updating the AllWells_updated layer on to ArcGIS Online.

```{r make spatial}
#-- create directory to store updated file
dir.create("./data/CalGEMwell_updates")

#-- specify filename
Updated.File = paste0("Well_updates_", Sys.Date())
Updated.File

#-- create a shapefile from new.wells.fmt
coords<-data.frame(x =  new.wells.fmt$Longitd , y = new.wells.fmt$Latitud  ) # set coordinates
crs<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"   # set projection
new.wells.spdf = SpatialPointsDataFrame(coords=coords, data =new.wells.fmt,  proj4string = CRS(crs))  # create spdf
writeOGR(new.wells.spdf, "./data/CalGEMwell_updates", Updated.File, driver="ESRI Shapefile")  # write shapefile
# cgem = readOGR("./data/CalGEMwell_updates", Updated.File) # read in file

```

## Next Steps

A new shapefile has been created with the CalGEM updates. This file will be used to update the AllWells_updated layer. Follow instructions here: <https://support.esri.com/en/Technical-Article/000015842>
